Are parental care induced gene expression patterns affect more by internal cues or external stimuli?

### 2019-09-06 docs/characterization_manuscript.tex
two manuscripts not one

\newcommand{\getAff}[1]{
  \getIndex{#1}{Some Institute of Technology,Another University}
}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Alice Anonymous\textsuperscript{\getAff{Some Institute of Technology}}\textsuperscript{*},
Bob Security\textsuperscript{\getAff{Another University}}\\
\bigskip
\textbf{\getAff{Some Institute of Technology}}Department, Street, City, State, Zip\\
\textbf{\getAff{Another University}}Department, Street, City, State, Zip\\
\bigskip
* Corresponding author: alice@example.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
One of the central questions of animal behavior is how biological
mechanisms mediate and respond to behavior. Successful rearing of
offspring involves a shift from aggressive and sexual behaviors to more
caring and nurturing ones, but what neural substrates mediate this
transition? How flexible are these changes in response to unpredictable
environmental perturbations, and how is behavior altered because of
them? In a rapidly changing world, understanding how the environment
affects the brain and how, in turn, the brain affects the behavioral
transition into parental care will shed light on how changes in
environment can ultimately affect fitness. Previous discoveries have
implicated particular hormones that play a role in the maintenance of
parental care behavior in vertebrates, such as: oxytocin and vasopressin
(avian homologs: mesotocin and vasotocin), vasoactive intestinal
peptide, and prolactin. However, we know very little about when and how
the brain transitions into parental care behaviors in any vertebrate,
and this knowledge is fundamentally important to our understanding of
the mechanisms mediating parental care. Here, we propose to expand our
knowledge by using classic offspring replacement and removal
manipulations to uncover the role of GnIH during the transition to
parental care behaviors. In addition to this targeted approach, we
propose a powerful untargeted approach to explore what other changes in
the brain may be occurring during this transition. We will use
high-throughput sequencing technology and immunohistochemistry to
uncover all genetic changes and specific proteomic changes in the brain,
along with their relationship with GnIH, which occur during the
transition to parental care. Our long-term goal is to understand the
interplay of genes, proteins, and parental care. Specifically, we aim to
understand the genetic and physiological mechanisms driving parental
care behaviors in all vertebrates, and to document this knowledge in a
rich, open database. This knowledge would enable manipulation of
specific parental care behaviors and their activation using emerging
powerful genetic technologies.

@@

This ecludian clustering shows many clusters are heterogenous,
consisting of sames from all tissues types, both sexes, and from
multiple treatment groups.

\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}

\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
hypothalamus and pituitary samples separte form the gonad samples, but
some are clusters contain samples from all tissues.The top panel is a
dendrogram showing Euclidean distance of sample. Females (white), males
(red), gonad (white), hypothalamus (pink), pituitary (red).

The rest of the analyses are peformed on each sex and tissue group
separately, a 2 x 3 design.

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}

@@

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-5-1.pdf}

\includegraphics{../figures/characterization/musicalgenes-05.png}

\hypertarget{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}{%
\subsection{Elucidate cause of transcription \& protein changes during
parental care
transitions}\label{elucidate-cause-of-transcription-protein-changes-during-parental-care-transitions}}

Using egg and chick removal and replacement experiments we will
determine if neural transcription and translation during parental care
transitions are based on external sensory information (the
presence/absence of eggs or chicks) or on an internal clock-timing
mechanism (natural biological rhythms unaffected by changes in external
sensory information). As rock doves are known to cross-foster eggs and
chicks, our working hypothesis is that the transition to parental care
behaviors is based on external sensory input, and specific predictions
are explained in conjunction with each manipulation below.

\hypertarget{give-birds-eggs}{%
\subsubsection{Give birds eggs}\label{give-birds-eggs}}

We will place eggs in nests of birds lacking them and sample brains the
following day. Prediction: birds will transition into incubation
behavior, and neural transcription and translation will mirror that of
birds sampled on the day their first egg was laid. This would support
our working hypothesis that sensory perception of the environment drives
these changes. If we do not observe changes in behavior, transcription,
and translation, or they only occur in males, one of many
interpretations is that males respond to environmental cues while
females are cued by an internal clock-timing mechanism regulating
transitions into parental care.

\hypertarget{prolong-incubation}{%
\subsubsection{Prolong incubation}\label{prolong-incubation}}

We will prolong incubation by removing eggs from actively incubated
nests on the day before hatching (Day 17) and immediately replace them
with infertile, `dummy' eggs. We will then sample brains 3 days after
normal hatch time (Day 21). Collaborator Silver and colleagues have
shown sex differences in the endocrine profiles of male and female doves
in response to prolonged incubation, in that prolactin levels are
maintained in incubating females but not males (Ramsey et al.~1985).
Because of this, we predict that if changes in behavior, neural
transcription, and translation occur in females because of sensory
information related to chicks hatching (supporting our working
hypothesis), then females that experience this prolonged incubation will
mirror the changes of incubating birds sampled on Day 17 of incubation.
However, if changes in behavior, transcription and translation are
brought on by internal clock mechanisms rather than external sensory
information, as predicted at this stage in males, bird behavior and
biology will resemble that of birds collected on Day 3 post-hatching.

\hypertarget{extend-time-of-hatching}{%
\subsubsection{Extend time of hatching}\label{extend-time-of-hatching}}

We will prolong incubation by removing eggs from nests undergoing active
incubation on Day 17 (the end of incubation) and immediately replacing
eggs with `dummy' eggs. On Day 21 (3 days post-natural hatch time) we
will offer hatchlings. Doves are known to care for nestlings placed in
their nest that are not their own (Klinghammer and Hess 1964, Hasen
1971). Brains will be sampled the following day. Prediction: Behavior,
transcription and translation will be similar to birds naturally caring
for chick(s) one day post-hatch of the first chick, supporting our
working hypothesis. Alternatively, if behavior, transcription and
translation are similar to birds caring for hatchlings on Day 3
post-hatching, which would have been the normal time course for parental
care if chicks had hatched naturally, this would support regulation by
an internal clock mechanism.

\hypertarget{remove-eggs-or-chicks}{%
\subsubsection{Remove eggs or chicks}\label{remove-eggs-or-chicks}}

We will remove eggs during incubation on Day 1 (beginning), Day 9
(middle), or Day 17 (end) and sample brains on the following day. Note:
This is not a repeated measures sampling, and three groups of
male-female pairs will be independently sampled at each time point.
Prediction: Birds will revert to a behavioral, transcriptional and
translational pattern characteristic of the ``Nesting, prior to lay''
sampling point, lending support to our working hypothesis. If birds
maintain a profile characteristic of birds collected at similar time
points whose eggs have not been removed, this lends support to an
internal clock mechanism.

We will remove newly, naturally hatched chick(s) on Day 2: second chick
hatches and sample brains from parents the following day (Day 2 post
(second) egg hatching). Prediction: Behavior, transcription and
translation are sensory driven and will revert to a pattern
characteristic of the `Nesting, prior to lay' point, supporting our
working hypothesis. Alternatively, if an internal clock mechanism is at
play, we would expect to see a pattern similar to birds caring for
chicks on Day 2 post (second) egg hatching.

\includegraphics{../figures/sexes/offspringremoval-1.png}

\hypertarget{reduce-time-to-hatching}{%
\subsubsection{Reduce time to hatching}\label{reduce-time-to-hatching}}

We will remove eggs on Day 8 (middle) of incubation and immediately
replace them with newly hatched chicks. We will sample brains on the
following day (Day 9). Prediction: Changes in behavior, transcription
and translation are externally sensory driven, and these will be similar
to birds sampled on `Day 1: 1st egg hatches, supporting our working
hypothesis. Alternatively, if profiles remain similar to birds sampled
on Day 9 that are actively incubating eggs, this would support the
presence of an internal clock mechanism.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus
finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
pulvinar, volutpat risus.

\hypertarget{supplementary-figures}{%
\section{Supplementary Figures}\label{supplementary-figures}}

\includegraphics{manuscript_draft_files/figure-latex/wgcna-female-1.pdf}

\textbf{Fig S1a. Female characterization dendrogram with trait
heatmap.}. Most pituitary samples form a single cluster. More
hypothalamus samples cluster with gonads than with pituitary. The top
panel is a dendrogram showing Euclidean distance of sample. Females
(white), males (red), gonad (white), hypothalamus (pink), pituitary
(red).

\includegraphics{manuscript_draft_files/figure-latex/wgcna-male-1.pdf}

\textbf{Fig S1b. Male characterization dendrogram with trait heatmap.}.
As with females, most pituitary samples form a single cluster, while
more hypothalamus samples cluster with gonads than with pituitary.

>>>>

\newcommand{\getAff}[1]{
  \getIndex{#1}{UC Davis,UNH}
}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Peaks and valleys of prolatin-related gene expression during pigeon
parental care stages} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rayna M Harris\textsuperscript{\getAff{UC Davis}}\textsuperscript{*},
Suzanne Austin\textsuperscript{\getAff{UC Davis}},
Andrew Lang\textsuperscript{\getAff{UNH}},
Matthew MacManes\textsuperscript{\getAff{UNH}},
Rebecca M Calisi\textsuperscript{\getAff{UC Davis}}\\
\bigskip
\textbf{\getAff{UC Davis}}UC Davis, Davis, CA\\
\textbf{\getAff{UNH}}UNH, smalltown, NH\\
\bigskip
* Corresponding author: rmharris@ucdavis\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit.

@@

\includegraphics{characterization_manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}

@@

### 2019-09-06 docs/manipulation_manuscript.Rmd
two manuscripts not one

---
title: Title of submission to PLOS journal
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    corresponding: alice@example.com
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
author_summary: |
  One of the central questions of animal behavior is how biological mechanisms mediate and respond to behavior. Successful rearing of offspring involves a shift from aggressive and sexual behaviors to more caring and nurturing ones, but what neural substrates mediate this transition? How flexible are these changes in response to unpredictable environmental perturbations, and how is behavior altered because of them? In a rapidly changing world, understanding how the environment affects the brain and how, in turn, the brain affects the behavioral transition into parental care will shed light on how changes in environment can ultimately affect fitness. Previous discoveries have implicated particular hormones that play a role in the maintenance of parental care behavior in vertebrates, such as: oxytocin and vasopressin (avian homologs: mesotocin and vasotocin), vasoactive intestinal peptide, and prolactin. However, we know very little about when and how the brain transitions into parental care behaviors in any vertebrate, and this knowledge is fundamentally important to our understanding of the mechanisms mediating parental care. Here, we propose to expand our knowledge by using classic offspring replacement and removal manipulations to uncover the role of GnIH during the transition to parental care behaviors. In addition to this targeted approach, we propose a powerful untargeted approach to explore what other changes in the brain may be occurring during this transition. We will use high-throughput sequencing technology and immunohistochemistry to uncover all genetic changes and specific proteomic changes in the brain, along with their relationship with GnIH, which occur during the transition to parental care. Our long-term goal is to understand the interplay of genes, proteins, and parental care. Specifically, we aim to understand the genetic and physiological mechanisms driving parental care behaviors in all vertebrates, and to document this knowledge in a rich, open database. This knowledge would enable manipulation of specific parental care behaviors and their activation using emerging powerful genetic technologies.

bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
---

@@

Understanding the mechanisms underlying parental care are critical to circumventing issues with parent-newborn bonding as well, where ultimate explanations are obvious, but specific mechanisms remain elusive. The rock dove (_Columba livia_) is an ideal system to characterize changes in genetic expression during parental care transitions because: 1) ample genomic resources are available (Gillespie et al. 2013), including a complete annotated genome assembly @Shapiro1063 and methodology concerning reproductive physiology and behavior (Dong et al. 2012); and 2) rock doves are prolific, year-round breeders that thrive in captivity, making observation, manipulation, and sampling highly feasible year-round.

```{r, out.width = "360px", echo = F}
knitr::include_graphics("../figures/images/PigeonHPGaxis.png")
```

Rock doves are socially monogamous and offer bi-parental care, making inter- and intra-sexual comparisons possible. Birds offer two important behavioral transition points into parental care: the incubation of eggs and the caring for chicks. This produces two unique opportunities to study how the brain transitions into two different suites of parental care behaviors. Additionally, rock doves exhibit a parental care strategy analogous to mammals in that they, too, ‘lactate’ to feed their young (Gillespie et al. 2011, 2012). This lactation, unlike simple regurgitation of food, consists of the production and sloughing off of skin cells inside the crop sac of females and males, creating a protein-rich milk-like substance on which they rear their chicks. Many functional similarities between rock dove and mammalian lactation exist concerning the mediation of this event by the hormone prolactin (Dumont 1965). Additionally, like mammalian milk, rock dove milk delivers essential immunoglobulins and nutritional benefits to young, aiding in their immune function and development of microbiota @Gillespie2012. Thus, because rock doves incubate eggs and exhibit mammalian-like mediation and function of lactation for young, they have the potential to serve as a powerful theoretical bridge to understand the neurobiology of both avian and mammalian transitions into parental care.

Aim I lays the necessary groundwork for establishing a pattern of correlative relationships between observed changes in gene expression and patterns of parental behavior. With this pattern established, we can proceed with experimental manipulations proposed under Aim III to determine causation.

@@

```{r, out.width = "360px", echo = F}
knitr::include_graphics("../figures/images/samplingtimepoints.png")
```

## Characterize changes in neural-gene expression during parental care transitions

Our working hypothesis is that distinct changes in transcription occur in the brain at the anticipation of, during, and in response to two different types of parental care: incubation behavior and hatchling care.

This ecludian clustering shows many clusters are heterogenous, consisting of sames from all tissues types, both sexes, and from multiple treatment groups.

```{r wgcna, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
source("../analysis/08_WGCNA_1.R")
subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary",
                           "male.gonad", "male.hypothalamus", "male.pituitary"),
            mytitle = "Characterization dendrogram of samples")
```

**Fig 1. Characterization dendrogram with trait heatmap.**.
Most hypothalamus and pituitary samples separte form the gonad samples, but some are clusters contain samples from all tissues.The top panel is a dendrogram showing Euclidean distance of sample. Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).

The rest of the analyses are peformed on each sex and tissue group separately, a 2 x 3 design.

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
library(cowplot)
ggdraw() + draw_image("../figures/sexes/characterization-2.png") 
```

**Fig 2. The magnitude of gene expression changes between each parental transition**.
4-5K genes are differentially expressed beteen control birds and their nest buliding conspecifics in all tissues except the male gonad. 500 - 1000 genes are differentially expressed in the female pituitary from mid-late incubation as well as in the female hypothalamus and pituitary from hatch to nestingly care day 5.

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
a <- ggdraw() + draw_image("../figures/characterization/pca-2.png") 
b <- ggdraw() + draw_image("../figures/characterization/pca-5.png")

plot_grid(a,b)
```

![](../figures/characterization/musicalgenes-05.png)

@@

![](../figures/sexes/offspringremoval-1.png)

@@

```{r wgcna-female, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
subsetWGCNA(whichgroups = c("female.gonad", "female.hypothalamus", "female.pituitary"),
            mytitle = "Female sample dendrogram with trait heatmap")
```

**Fig S1a. Female characterization dendrogram with trait heatmap.**. 
Most pituitary samples form a single cluster. More hypothalamus samples cluster with gonads than with pituitary. The top panel is a dendrogram showing Euclidean distance of sample. Females (white), males (red), gonad (white), hypothalamus (pink), pituitary (red).

```{r wgcna-male, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
subsetWGCNA(c("male.gonad", "male.hypothalamus", "male.pituitary"),
            mytitle = "Male sample dendrogram with trait heatmap")
```

**Fig S1b. Male characterization dendrogram with trait heatmap.**. 
 As with females, most pituitary samples form a single cluster, while more hypothalamus samples cluster with gonads than with pituitary.

>>>>

---
title: Are parental care induced gene expression patterns affect more by internal cues or external stimuli?
author:
  - name: Rayna M Harris
    email: rmharris@ucdavis
    affiliation: UC Davis
    corresponding: "rmharris@ucdavis"
  - name: Suzanne Austin
    affiliation: UC Davis
  - name: Andrew Lang
    affiliation: UNH
  - name: Matthew MacManes
    affiliation: UNH
  - name: Rebecca M Calisi
    affiliation: UC Davis           
address:
  - code: UC Davis
    address: UC Davis, Davis, CA
  - code: UNH
    address: UNH, smalltown, NH
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. 
bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
---

@@

@@

@@

```{r, message=FALSE, warning=FALSE, cache=TRUE, echo = F, fig.width= 5}
library(cowplot)
ggdraw() + draw_image("../figures/manipulation/offspringremoval-1.png") 
```

@@

### 2019-09-06 docs/manipulation_manuscript.tex
two manuscripts not one

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Title of submission to PLOS journal} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Alice Anonymous\textsuperscript{\getAff{Some Institute of Technology}}\textsuperscript{*},
Bob Security\textsuperscript{\getAff{Another University}}\\
\bigskip
\textbf{\getAff{Some Institute of Technology}}Department, Street, City, State, Zip\\
\textbf{\getAff{Another University}}Department, Street, City, State, Zip\\
\bigskip
* Corresponding author: alice@example.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
One of the central questions of animal behavior is how biological
mechanisms mediate and respond to behavior. Successful rearing of
offspring involves a shift from aggressive and sexual behaviors to more
caring and nurturing ones, but what neural substrates mediate this
transition? How flexible are these changes in response to unpredictable
environmental perturbations, and how is behavior altered because of
them? In a rapidly changing world, understanding how the environment
affects the brain and how, in turn, the brain affects the behavioral
transition into parental care will shed light on how changes in
environment can ultimately affect fitness. Previous discoveries have
implicated particular hormones that play a role in the maintenance of
parental care behavior in vertebrates, such as: oxytocin and vasopressin
(avian homologs: mesotocin and vasotocin), vasoactive intestinal
peptide, and prolactin. However, we know very little about when and how
the brain transitions into parental care behaviors in any vertebrate,
and this knowledge is fundamentally important to our understanding of
the mechanisms mediating parental care. Here, we propose to expand our
knowledge by using classic offspring replacement and removal
manipulations to uncover the role of GnIH during the transition to
parental care behaviors. In addition to this targeted approach, we
propose a powerful untargeted approach to explore what other changes in
the brain may be occurring during this transition. We will use
high-throughput sequencing technology and immunohistochemistry to
uncover all genetic changes and specific proteomic changes in the brain,
along with their relationship with GnIH, which occur during the
transition to parental care. Our long-term goal is to understand the
interplay of genes, proteins, and parental care. Specifically, we aim to
understand the genetic and physiological mechanisms driving parental
care behaviors in all vertebrates, and to document this knowledge in a
rich, open database. This knowledge would enable manipulation of
specific parental care behaviors and their activation using emerging
powerful genetic technologies.

@@

Understanding the mechanisms underlying parental care are critical to
circumventing issues with parent-newborn bonding as well, where ultimate
explanations are obvious, but specific mechanisms remain elusive. The
rock dove (\emph{Columba livia}) is an ideal system to characterize
changes in genetic expression during parental care transitions because:
1) ample genomic resources are available (Gillespie et al.~2013),
including a complete annotated genome assembly {[}1{]} and methodology
concerning reproductive physiology and behavior (Dong et al.~2012); and
2) rock doves are prolific, year-round breeders that thrive in
captivity, making observation, manipulation, and sampling highly
feasible year-round.

\includegraphics[width=360px]{../figures/images/PigeonHPGaxis}

Rock doves are socially monogamous and offer bi-parental care, making
inter- and intra-sexual comparisons possible. Birds offer two important
behavioral transition points into parental care: the incubation of eggs
and the caring for chicks. This produces two unique opportunities to
study how the brain transitions into two different suites of parental
care behaviors. Additionally, rock doves exhibit a parental care
strategy analogous to mammals in that they, too, `lactate' to feed their
young (Gillespie et al.~2011, 2012). This lactation, unlike simple
regurgitation of food, consists of the production and sloughing off of
skin cells inside the crop sac of females and males, creating a
protein-rich milk-like substance on which they rear their chicks. Many
functional similarities between rock dove and mammalian lactation exist
concerning the mediation of this event by the hormone prolactin (Dumont
1965). Additionally, like mammalian milk, rock dove milk delivers
essential immunoglobulins and nutritional benefits to young, aiding in
their immune function and development of microbiota {[}2{]}. Thus,
because rock doves incubate eggs and exhibit mammalian-like mediation
and function of lactation for young, they have the potential to serve as
a powerful theoretical bridge to understand the neurobiology of both
avian and mammalian transitions into parental care.

Aim I lays the necessary groundwork for establishing a pattern of
correlative relationships between observed changes in gene expression
and patterns of parental behavior. With this pattern established, we can
proceed with experimental manipulations proposed under Aim III to
determine causation.

@@

\includegraphics[width=360px]{../figures/images/samplingtimepoints}

\hypertarget{characterize-changes-in-neural-gene-expression-during-parental-care-transitions}{%
\subsection{Characterize changes in neural-gene expression during
parental care
transitions}\label{characterize-changes-in-neural-gene-expression-during-parental-care-transitions}}

Our working hypothesis is that distinct changes in transcription occur
in the brain at the anticipation of, during, and in response to two
different types of parental care: incubation behavior and hatchling
care.

This ecludian clustering shows many clusters are heterogenous,
consisting of sames from all tissues types, both sexes, and from
multiple treatment groups.

\includegraphics{manuscript_draft_files/figure-latex/wgcna-1.pdf}

\textbf{Fig 1. Characterization dendrogram with trait heatmap.}. Most
hypothalamus and pituitary samples separte form the gonad samples, but
some are clusters contain samples from all tissues.The top panel is a
dendrogram showing Euclidean distance of sample. Females (white), males
(red), gonad (white), hypothalamus (pink), pituitary (red).

The rest of the analyses are peformed on each sex and tissue group
separately, a 2 x 3 design.

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}

\textbf{Fig 2. The magnitude of gene expression changes between each
parental transition}. 4-5K genes are differentially expressed beteen
control birds and their nest buliding conspecifics in all tissues except
the male gonad. 500 - 1000 genes are differentially expressed in the
female pituitary from mid-late incubation as well as in the female
hypothalamus and pituitary from hatch to nestingly care day 5.

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-5-1.pdf}

\includegraphics{../figures/characterization/musicalgenes-05.png}

@@

\includegraphics{../figures/sexes/offspringremoval-1.png}

@@

\includegraphics{manuscript_draft_files/figure-latex/wgcna-female-1.pdf}

\textbf{Fig S1a. Female characterization dendrogram with trait
heatmap.}. Most pituitary samples form a single cluster. More
hypothalamus samples cluster with gonads than with pituitary. The top
panel is a dendrogram showing Euclidean distance of sample. Females
(white), males (red), gonad (white), hypothalamus (pink), pituitary
(red).

\includegraphics{manuscript_draft_files/figure-latex/wgcna-male-1.pdf}

\textbf{Fig S1b. Male characterization dendrogram with trait heatmap.}.
As with females, most pituitary samples form a single cluster, while
more hypothalamus samples cluster with gonads than with pituitary.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Shapiro1063}{}%
1. Shapiro MD, Kronenberg Z, Li C, Domyan ET, Pan H, Campbell M, et al.
Genomic diversity and evolution of the head crest in the rock pigeon.
Science. American Association for the Advancement of Science; 2013;339:
1063--1067.
doi:\href{https://doi.org/10.1126/science.1230422}{10.1126/science.1230422}

\leavevmode\hypertarget{ref-Gillespie2012}{}%
2. Gillespie MJ, Stanley D, Chen H, Donald JA, Nicholas KR, Moore RJ, et
al. Functional Similarities between Pigeon `Milk' and Mammalian Milk:
Induction of Immune Gene Expression and Modification of the Microbiota.
Salmon H, editor. PLoS ONE. Public Library of Science; 2012;7: e48363.
doi:\href{https://doi.org/10.1371/journal.pone.0048363}{10.1371/journal.pone.0048363}

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Are parental care induced gene expression patterns affect more by
internal cues or external stimuli?} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rayna M Harris\textsuperscript{\getAff{UC Davis}}\textsuperscript{*},
Suzanne Austin\textsuperscript{\getAff{UC Davis}},
Andrew Lang\textsuperscript{\getAff{UNH}},
Matthew MacManes\textsuperscript{\getAff{UNH}},
Rebecca M Calisi\textsuperscript{\getAff{UC Davis}}\\
\bigskip
\textbf{\getAff{UC Davis}}UC Davis, Davis, CA\\
\textbf{\getAff{UNH}}UNH, smalltown, NH\\
\bigskip
* Corresponding author: rmharris@ucdavis\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit.

@@

@@

@@

\includegraphics{manipulation_manuscript_files/figure-latex/unnamed-chunk-2-1.pdf}

@@

@@
